---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

Loading required libraries
```{r library, echo = TRUE}
library(knitr)
library(dplyr)
library(ggplot2)
```

## Loading and preprocessing the data
```{r loading, echo=TRUE}
setwd("/Users/silencemac/Desktop/coursera/course5.2")
activity <- read.csv("activity.csv", sep = ",")
activity$date <- as.Date(activity$date)
```

## What is mean total number of steps taken per day?
```{r sum, echo=TRUE}
byday <- aggregate(steps ~ date, activity, sum)
hist(byday$steps, xlab = "steps", ylab = "frequency", 
     main = "total number of steps per day")
mean(byday$steps)
median(byday$steps)
```

## What is the average daily activity pattern?

```{r mean, echo=TRUE}
byinterval <- aggregate(steps ~ interval, activity, mean)
plot(byinterval$interval, byinterval$steps, type = "l", xlab = "interval",
     ylab = "average steps", main = "Average steps over all day")
```

Which 5-minute interval contains the maximum number of steps?
```{r max, echo=TRUE}
maxsteps <- which.max(byinterval$steps)
byinterval[maxsteps,]
```

## Imputing missing values
```{r NA, echo=TRUE}
NAsum <- sum(is.na(activity))
paste("The number of missing values in the dataset is", NAsum)
```

Devise a strategy for filling in all of the missing values in the dataset
:first to figure out the "NA", then replace the NA with the mean of steps by intervals
```{r fill, echo=TRUE}
fillact <- activity
NAact <- is.na(fillact$steps)
mean.interval <- tapply(fillact$steps, fillact$interval, mean, 
                        na.rm = TRUE, simplify = TRUE)
```

Create a new dataset with the missing data filled in
```{r, echo=TRUE}
fillact$steps[NAact] <- mean.interval[as.character(fillact$interval[NAact])]
```

Make a histogram of the total number of steps taken each day
```{r plot1, echo=TRUE}
byday.NA <- aggregate(steps ~ date, fillact, sum)
hist(byday.NA$steps, xlab = "steps", ylab = "frequency", 
     main = "Total number of steps per day")
m2 <- mean(byday.NA$steps)
md2 <- median(byday.NA$steps)
m1 <- mean(byday$steps)
md1 <- median(byday$steps)
paste("The mean and median total number of steps taken per day are", m2, "and", md2)
paste("In the first part, the mean and median total number of steps taken per day are", 
      m1, "and", md1)
```

## Are there differences in activity patterns between weekdays and weekends?
```{r days, echo=TRUE}
fillact$date2 <- weekdays(fillact$date)
fillact$days[fillact$date2 %in% c("Saturday", "Sunday")] <- "weekend"
fillact$days[fillact$date2 %in% c("Monday", "Tuesday", "Wednesday", 
                                 "Thursday", "Friday")] <- "weekday"
byinteral_days <- aggregate(steps ~ interval + days, fillact, mean)
g <- ggplot(byinteral_days, aes(interval, steps, color=days))
p <- g + geom_line() + labs(x = "intervals", y = "average steps", 
                            title = "Average steps on weekdays and weekend") + 
       facet_wrap(~ days, ncol = 1)
print(p)
```
